Learning to Synthesize Rules for Rule-based Intrusion Detectors

### 2020-03-06 paper/paper.tex
pass on pending fixes

The content strategy focuses on extracting content options from the
message payload. By inspecting existing rules, we found that these
options refer to sequences of characters separated by typical natural
language delimiters such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. To find the values associated with these
options, \tname{} splits the payload in tokens using these delimiters.
Considering the example from Section~\ref{sec:content-example},
\tname{} produced a total of 23 tokens\Mar{@Lucas, isto estah certo?
  Se sim, como isto cai para 5 (pela Tabela 3)} \Luc{6.1.1 eh o ataque 
JBoss. Na tabela constam 23 tokens.}, considered content
option candidates.

It is worth noting that we found that deciding which reverse
engineering strategy to use based on the protocol of the input
(malicious) messages was effective. We used the payload strategy for
http and the field strategy for the other protocols. For flooding
attacks, \tname\ runs the same approach for each message and
identifies common options across all messages. For that case,
\tname\ reports, in addition, a \CodeIn{threshold} option, as
discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
Reverse Engineering step, \tname\ produces a seed rule that captures
the malicious traffic.

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
that is passed as parameter\Mar{isto realmente eh necessario? se sim,
  explique. me pareceu que o que vc. precisa ate aqui jah estaria no
  log do \suri}, 2)~replaying the negative benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. \Mar{discuss
  filter function.} It is worth
noting that the benign traffic is stored in \emph{pcap}
files~\cite{pcap} and Suricata can process those files very
efficiently compared to analyzing network messages directly. Instead
of sending traffic to a machine with a deployed server (\eg{}, web
server) that \suri\ would monitor, our setup calls \suri\ directly
through an API. \suri\ is invoked multiple times in the process of
minimizing the rule.  \textbf{Limitations.} It is worth noting that
the performance of \tname{} depends on the quality of inputs. For the
positive malicious traffic, \tname{} only requires one sample (as
input to step one), but can benefit of several samples (see
line~\ref{filtering} of Algorithm~\ref{algo1}). The same applies to
the negative traffic. However, in this case, \tname{} can use a large
variety of public data sets available online (see
Section~\ref{sec:objectanalysis-benign}).

@@

\subsubsection{Weight Optimization}
\label{sec:weight-optimization}
We optimized the weights of these heuristic functions as follows. We
ran \tname{} using different combinations of weights on a sample of
attacks and chose the combination that offered the best overall
performance, which we computed with the sum of the rankings of the
golden rule on each attack. The weight combination that led to the
smaller sum of positions (of the golden rule) in the rakings was
considered the best choice. We considered \emph{all} combinations of
the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
four different functions and five possible weight values, we needed to
run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
end of the process, the weight combination $\langle{}x,y,z,w\rangle$
was the best we found. \Mar{$\leftarrow$indique aqui quais foram os
  pesos que a gente usar para cada funcao.} \Luc{f1: 1, f2: 0.25, f3: 0.25, f4: 1} Although empirical results
showed that this strategy results in good performance, we remain to
evaluate other optimization strategies, such as evolutionary
alternatives as AVM~\cite{10.1109/32.57624}, to find better weights.

>>>>

The content strategy focuses on extracting content options from the
message payload. By inspecting existing rules, we found that these
options refer to sequences of characters separated by typical natural
language delimiters such as $\backslash$t, $\backslash$n,
$\backslash$r, and spaces. To find the values associated with these
options, \tname{} splits the payload in tokens using these delimiters.

%Considering the example from Section~\ref{sec:active-recon},
%\tname{} produced a total of 5 tokens.

It is worth noting that we found that deciding which reverse
engineering strategy to use based on the protocol of the input
(malicious) messages showed to be effective. We used the payload
strategy for http and the field strategy for the other protocols. For
flooding attacks, \tname\ runs the same approach for each message and
identifies common options across all messages. For that case,
\tname\ reports, in addition, a \CodeIn{threshold} option, as
discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
Reverse Engineering step, \tname\ produces a seed rule that captures
the malicious traffic.

@@

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~spawning Suricata to monitor a specified
rule, 2)~replaying the negative benign traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. The function
\emph{filter} is implemented similarly. It returns a projection of the
set $\mathcal O$ where every rule captures the attacks characterized
by $\mathit{mTr}$. It is worth noting that Suricata can process
\emph{pcap} files~\cite{pcap} very efficiently compared to analyzing
network messages directly. Instead of sending traffic to a machine
with a deployed server (\eg{}, web server) that \suri\ would monitor,
our setup calls \suri\ directly through an API. \suri\ is invoked
multiple times for creating rules.

%% \textbf{Limitations.} It is worth
%% noting that the performance of \tname{} depends on the quality of
%% inputs. For the positive malicious traffic, \tname{} only requires one
%% sample (as input to step one), but can benefit of several samples (see
%% line~\ref{filtering} of Algorithm~\ref{algo1}). The same applies to
%% the negative traffic. However, in this case, \tname{} can use a large
%% variety of public data sets available online (see
%% Section~\ref{sec:objectanalysis-benign}).

@@

\subsubsection{Weight Optimization}
\label{sec:weight-optimization}
We optimized the weights of these heuristic functions as follows. We
ran \tname{} using different combinations of weights on a sample of
attacks and chose the combination that offered the best overall
performance, which we computed with the sum of the rankings of the
golden rule on each attack. The weight combination that led to the
smaller sum of positions (of the golden rule) in the rakings was
considered the best choice. We considered \emph{all} combinations of
the following weights 0, 0.25, 0.5, 0.75, and 1. Given that we have
four different functions and five possible weight values, we needed to
run \tname{} on the sample of attacks for 625 (=$5^4$) times. In the
end of the process, the weight combination
$\langle{}1,0.25,0.25,1\rangle$ was the best choice we found. Although
results showed that this strategy results in good performance, we
remain to evaluate other optimization strategies to find better
weights.

%% , such as evolutionary alternatives as
%% AVM~\cite{10.1109/32.57624}.
